Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-22 paper.tex

The compression was performed on a single compute node with 64 CPUs and 416 GiB of RAM.  
We first converted the VCF to an ICF, which took 6 hours and 17 minutes of  
effective runtime and utilized 343 hours of CPU time, achieving a parallel efficiency of 86\%.  
This step incurred a cost of \$23.88. The resulting ICF representation required 7.5 GiB of storage  
distributed across 91,232 files. Next, the ICF was converted to Zarr, which took 50 minutes and 32 
seconds of effective runtime and 49 hours and 49 minutes of CPU time, with a parallel efficiency of 92\%, 
at a cost of \$3.19. The final Zarr representation occupied 13 GiB of storage across 369,656 chunk files.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced 
the from the 245,388 exome-like samples from chromosome 20
using \texttt{vcf2zarr} and chunk sizes of 1,000 variants and 
10,000 samples.
For column details see the caption of Table~\ref{tab-genomics-england-data}.
\label{tab-all-of-us-data}}
\begin{tabular}{llS[table-format=3.2]S[table-format=3.2]S[table-format=3.2]}
\toprule
{Field} & {type} & {storage} & {compress} & {\%total} \\
\midrule
/call\_GQ & int8 & 1.89G & 87.0 & 45.94\% \\
/call\_genotype & int8 & 914.22M & 370.0 & 21.70\% \\
/call\_RGQ & int16 & 729.57M & 460.0 & 17.32\% \\
/call\_genotype\_mask & bool & 606.02M & 550.0 & 14.39\% \\
/call\_genotype\_phased & bool & 17.17M & 9700.0 & 0.41\% \\
/variant\_allele & str & 4.74M & 110.0 & 0.11\% \\
/variant\_filter & bool & 2.87M & 0.9 & 0.07\% \\
% /variant\_AN & int32 & 908.75 K & 3.1 & 0.02\% \\
% /variant\_position & int32 & 810.77 K & 3.4 & 0.02\% \\
% /sample\_id & object & 357.15 K & 5.4 & 0.01\% \\
% /variant\_length & int16 & 152.4 K & 9.2 & 0.00\% \\
% /variant\_id & object & 56.18 K & 99.0 & 0.00\% \\
% /variant\_quality & float32 & 51.92 K & 54.0 & 0.00\% \\
% /variant\_id\_mask & bool & 51.89 K & 13.0 & 0.00\% \\
% /variant\_contig & int16 & 50.6 K & 28.0 & 0.00\% \\
% /region\_index & int32 & 13.93 K & 1.2 & 0.00\% \\
% /contig\_length & int64 & 9.24 K & 2.8 & 0.00\% \\
% /contig\_id & object & 8.33 K & 3.2 & 0.00\% \\
% /filter\_id & object & 4.48 K & 0.0 & 0.00\% \\
\bottomrule
\end{tabular}
\end{table}

>>>>

The compression was performed on the All of Us Researcher Workbench platform 
(based on Google Cloud), using a single compute node with 64 CPUs and 416 GiB 
of RAM, which cost \$3.80/hour. We first converted the VCF to an ICF, which 
took 6 hours and 17 minutes of effective runtime and utilized 343 hours of CPU 
time, achieving a parallel efficiency of 86\%. This step incurred a cost of 
\$23.88. The resulting ICF representation required 7.5 GiB of storage distributed 
across 91,232 files. Next, the ICF was converted to Zarr, 
after removing the field \texttt{call\_FT}, which is 
suspected to be problematic. The conversion took 53 minutes and 40 seconds, with 
a CPU time of 52 hours and 34 minutes and a parallel efficiency of 92\%, at a 
cost of \$3.40. This second Zarr representation occupied 8 GiB of storage across 
214,190 chunk files.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced 
    from the dataset using \texttt{vcf2zarr} with chunk sizes of 1,000 variants and 
    10,000 samples. For column details, see Table~\ref{tab-genomics-england-data}.
    \label{tab-top-fields-data}}
\begin{tabular}{llS[table-format=3.2]S[table-format=4.1]S[table-format=3.2]}
\toprule
{Field} & {Type} & {Storage} & {Compress} & {\% Total} \\
\midrule
/call_LAD & int16 & 2.26 GiB & 290.0 & 29.63% \\
/call_GQ & int8 & 1.89 GiB & 87.0 & 24.78% \\
/call_LA & int8 & 1.24 GiB & 260.0 & 16.26% \\
/call_genotype & int8 & 914.22 MiB & 370.0 & 11.70% \\
/call_RGQ & int16 & 729.57 MiB & 460.0 & 9.34% \\
/call_genotype_mask & bool & 606.02 MiB & 550.0 & 7.76% \\
\bottomrule
\end{tabular}
\end{table}
